---
output:
  github_document:
    toc: false
    toc_depth: 3
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  echo = TRUE,
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(knitr)
library(devtools)
devtools::load_all()
devtools::document()
```

# gwkit

<!-- badges: start -->
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/ftsiboe/gwkit/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ftsiboe/gwkit/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/ftsiboe/gwkit/graph/badge.svg?token=MfnCHBGcET)](https://codecov.io/gh/ftsiboe/gwkit)
![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)
![R >= 4.1](https://img.shields.io/badge/R-%3E=4.1-blue)
<!-- badges: end -->

# Introduction

`gwkit` is a convenience toolkit of wrappers and reproducible workflows around
[`GWmodel`](https://cran.r-project.org/package=GWmodel) and related spatial
packages (`sf`, `terra`, `sp`) for geographically weighted analysis. It *wraps*,
rather than reimplements, the underlying estimators, so common tasks reduce to a
few uniform calls that are agnostic to the spatial-unit identifier.

**Disclaimer:** `gwkit` builds on other authors' packages; please cite `GWmodel`
and the other dependencies where appropriate. See the
[LICENSE](https://github.com/ftsiboe/gwkit/blob/main/LICENSE) file for details.

# Installation

`gwkit` can be installed directly from GitHub:

```{r installation, echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
# install.packages("remotes")
remotes::install_github("ftsiboe/gwkit")
```

It depends on `GWmodel`, `sf`, `terra`, and `sp`; the county workflow also uses
`urbnmapr`, installed from GitHub via the `Remotes` field.

# Supported functionalities

**Geographically weighted summary statistics** — local means, standard
deviations, and related statistics at points, polygons, or U.S. counties:

* `estimate_gwss_by_point()`
* `estimate_gwss_by_polygon()`
* `estimate_gwss_by_county()`

**Geographically weighted mean/variance regression** — a local regression of a
`response` on a `covariate`, returning the local slope on the mean and
(optionally) on the variance of the response, with local standard errors,
t-values, and p-values:

* `estimate_gwr_by_point()`
* `estimate_gwr_by_polygon()`

**Distance-metric presets** — Euclidean, Manhattan, Minkowski, Chebyshev, and
Great Circle:

* `gw_distance_metric_presets()`
* `gw_distance_metric_names()`
* `resolve_distance_metric()`

**Consensus across the kernel × distance-metric domain** — reduce results
estimated over a grid of GW settings to a single class per unit (the mode across
settings), with optional first-order **Queen contiguity** reconciliation that
widens the neighbourhood order to break ties:

* `gw_optimal_class_by_point()`
* `gw_optimal_class_by_polygon()`

# Examples

The distance-metric presets are self-contained:

```{r metrics}
library(gwkit)
gw_distance_metric_names()
```

## Example 1: Geographically weighted mean/variance regression at points

Fit a local regression of `response` on `covariate` (one row per
unit-observation), returning the local slope on the mean and on the variance:

```{r gwr_point, eval=FALSE}
est <- estimate_gwr_by_point(
  data      = observations,          # one row per unit-observation
  unit      = "grid_id",
  response  = "precipitation",
  covariate = "trend",
  coords    = c("longitude", "latitude"),
  kernel    = "bisquare",
  distance_metric = "Great Circle"
)
```

## Example 2: Optimal class across a grid of GW settings

Sweep several kernel × distance-metric settings, then pick the optimal class per
unit — the mode across settings, de-speckled by a first-order Queen contiguity
vote (ties widen the neighbourhood order):

```{r consensus, eval=FALSE}
optimal <- gw_optimal_class_by_point(
  class_dt  = stacked_settings,      # rows stacked over kernel x distance_metric
  unit_col  = "grid_id",
  class_col = "trend_class_05"
)
```

The polygon variants (`estimate_gwr_by_polygon()`,
`gw_optimal_class_by_polygon()`) take a `terra` `SpatVector` or `sf` layer and
derive centroids / Queen contiguity from the geometry.

# 📚 Citation

If you find it useful, please consider starring the repository and citing the
underlying `GWmodel` package alongside `gwkit`.

# 📬 Contact

Constructive feedback is highly appreciated, and collaborations using this
package are actively encouraged. Please reach out by sending emails to Francis
Tsiboe ([ftsiboe@hotmail.com](mailto:ftsiboe@hotmail.com)).
