---
output:
  github_document:
    toc: false
    toc_depth: 3
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  echo = TRUE,
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(knitr)
library(devtools)
devtools::load_all()
devtools::document()
```

# gwkit

<!-- badges: start -->
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/ftsiboe/gwkit/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ftsiboe/gwkit/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/ftsiboe/gwkit/graph/badge.svg?token=MfnCHBGcET)](https://codecov.io/gh/ftsiboe/gwkit)
![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)
![R >= 4.1](https://img.shields.io/badge/R-%3E=4.1-blue)
<!-- badges: end -->

# Overview

`gwkit` is a convenience toolkit of wrappers and reproducible workflows around
[`GWmodel`](https://cran.r-project.org/package=GWmodel) and related spatial
packages (`sf`, `terra`, `sp`) for geographically weighted analysis. It *wraps*,
rather than reimplements, the underlying estimators, so common tasks reduce to a
few uniform calls that are agnostic to the spatial-unit identifier.

**Disclaimer:** `gwkit` builds on other authors' packages; please cite `GWmodel`
and the other dependencies where appropriate. See the
[LICENSE](https://github.com/ftsiboe/gwkit/blob/main/LICENSE) file for details.

# Installation

`gwkit` can be installed directly from GitHub:

```{r installation, echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
# install.packages("remotes")
remotes::install_github("ftsiboe/gwkit")
```

It depends on `GWmodel`, `sf`, `terra`, and `sp`; the county workflow also uses
`urbnmapr`, installed from GitHub via the `Remotes` field.

# Supported functionalities

Each estimator is a single, spatial-unit-agnostic call. Point vs polygon mode is
**class-detected**: pass an `sf`/`SpatVector` `geometry` and units are reduced to
their point-on-surface (Queen contiguity from shared boundaries); otherwise the
`coords` columns of the table are used (lattice contiguity).

**Geographically weighted summary statistics** (`estimate_gwss()`) — local means,
standard deviations, and, for two or more variables, the local
covariance/correlation at each unit.

**Geographically weighted regression** (`estimate_gwr()`) — a local regression
returning the slope on the mean (and optionally the variance) of the response in
a LONG table. Supplying `panel`/`time` switches to a **fixed-effects** local fit
(Frisch–Waugh–Lovell within transform), so the slope is identified from
within-unit, over-time variation.

**Geographically weighted spatial lag** (`estimate_gwlag()`) — the geographically
weighted mean of a variable over each unit's neighbours (`include_self = FALSE`),
the neighbouring-exposure term used in availability metrics.

**Distance-metric presets** — Euclidean, Manhattan, Minkowski, Chebyshev, and
Great Circle, via `gw_distance_metric_presets()`, `gw_distance_metric_names()`,
and `resolve_distance_metric()`.

**Consensus across the kernel × distance-metric domain** — reduce results
estimated over a grid of GW settings to one per-unit summary: `gw_consensus_scalar()`
for continuous estimates (robust reducer + sign-agreement share) and
`gw_consensus_class()` for categorical classes (modal class + Queen-contiguity
vote), both with optional first-order **Queen contiguity** de-speckling.

# Examples

A worked, end-to-end study lives in
[`data-raw/examples`](https://github.com/ftsiboe/gwkit/tree/main/data-raw/examples).
Using the packaged `us_state_ag_census` dataset, it asks a two-part question:
does farmland trade at a **risk discount** — cheaper where production risk is
higher — and does the **Federal Crop Insurance Program (FCIP)** offset that
discount where participation is high? The local regression is

$$Y = a_0 + a_1\,r + a_2\,x + a_3\,(r\cdot x) + \text{(state FE)},$$

where `Y` is `log(ag_land_value_per_acre)`, the risk proxy `r` is the FCIP base
premium rate (`fcip_base_rate`), and participation `x` is `fcip_adoption`.
`a1` is the **risk discount** (expected `< 0`) and `a3` the
**FCIP mitigation** (expected `> 0`); geographically weighted estimation maps
where each holds. A live caveat runs through the examples: separating `a1` from
permanent between-state differences requires within-state, over-time variation in
risk, which fixed effects strip — example 06 shows how thin that variation is.
The packaged data is state-level (a methods demo); the substantive design is
parcel-level.

| Function | Role in the study |
|---|---|
| `estimate_gwr()` | local risk discount `a1` & FCIP mitigation `a3` (cross-section + FE panel) — [01_estimate_gwr](https://github.com/ftsiboe/gwkit/blob/main/data-raw/examples/01_estimate_gwr.md) |
| `estimate_gwlag()` | the surrounding risk & insurance environment (spatial lag) — [02_estimate_gwlag](https://github.com/ftsiboe/gwkit/blob/main/data-raw/examples/02_estimate_gwlag.md) |
| `estimate_gwss()` | where risk and value co-move (local correlation) — [03_estimate_gwss](https://github.com/ftsiboe/gwkit/blob/main/data-raw/examples/03_estimate_gwss.md) |
| `gw_consensus_scalar()` | is the discount spec-robust? (consensus `a1` + sign agreement) — [04_gw_consensus_scalar](https://github.com/ftsiboe/gwkit/blob/main/data-raw/examples/04_gw_consensus_scalar.md) |
| `gw_consensus_class()` | sign class of the discount (modal + Queen vote) — [05_gw_consensus_class](https://github.com/ftsiboe/gwkit/blob/main/data-raw/examples/05_gw_consensus_class.md) |
| `gw_distance_metric_presets()` / `resolve_distance_metric()` / `fixed_effect_model_data_prep()` | building blocks + the FE identification problem — [06_building_blocks](https://github.com/ftsiboe/gwkit/blob/main/data-raw/examples/06_building_blocks.md) |

The distance-metric presets are self-contained:

```{r metrics}
library(gwkit)
gw_distance_metric_names()
```

The flagship call fits the risk-discount model locally at every state, with
polygon geometry supplied directly (the interaction is a pre-built column so it
survives fixed-effects demeaning):

```{r gwr_example, eval=FALSE}
data(us_state_ag_census)
gwr <- estimate_gwr(
  data    = d_cs, unit = "state_code",
  formula = lv ~ base_rate + participation + rate_x_part,   # lv = log($/acre)
  geometry = states, poly_id = "state_code",
  distance_metric = "Great Circle", kernel = "bisquare",
  adaptive = TRUE, bw = 12, terms = "base_rate"    # a1 = local risk discount
)
```

# 📚 Citation

If you find it useful, please consider starring the repository and citing the
underlying `GWmodel` package alongside `gwkit`.

# 📬 Contact

Constructive feedback is highly appreciated, and collaborations using this
package are actively encouraged. Please reach out by sending emails to Francis
Tsiboe ([ftsiboe@hotmail.com](mailto:ftsiboe@hotmail.com)).
