---
title: "gwkit example 02 — `estimate_gwlag()`: the surrounding risk & insurance environment"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 5, dpi = 150)
```

A state's farmland is priced partly against its *regional* context — the risk
and insurance environment of its neighbours. `estimate_gwlag()` is a **spatial
lag**: the geographically weighted mean of a variable over a unit's *neighbours*
(`include_self = FALSE`). Here it summarises the surrounding FCIP base rate (the
regional risk climate) and surrounding participation (regional insurance
uptake) — the kind of neighbouring-exposure terms that enter an
availability/spillover specification.

```{r data}
library(data.table); library(sf); library(ggplot2)
source("_setup.R")     # loads gwkit (dev tree if present) + the ERS framework

data(us_state_ag_census)
d <- data.table::copy(us_state_ag_census)
d[, base_rate     := fcip_base_rate]
d[, participation := fcip_adoption]
d <- d[is.finite(base_rate) & is.finite(participation) &
         census_year == max(census_year)]

states <- urbnmapr::get_urbn_map("states", sf = TRUE)
states$state_code <- as.integer(states$state_fips)
states <- states[states$state_code %in% d$state_code, ]
```

```{r gwlag}
lagged <- estimate_gwlag(
  data = d, unit = "state_code",
  value_cols = c("base_rate", "participation"),
  geometry = states, poly_id = "state_code",
  distance_metric = "Great Circle", kernel = "bisquare",
  adaptive = TRUE, bw = 8, include_self = FALSE)   # strict neighbour mean

lagged[, state_code := as.integer(state_code)]
head(lagged)   # *_LM columns = neighbour-weighted means
```

```{r gwlag-map}
# own vs neighbour-weighted FCIP base rate (the regional risk climate)
m_own <- merge(states, d[, .(state_code, val = base_rate)], by = "state_code")
m_lag <- merge(states, lagged[, .(state_code, val = base_rate_LM)], by = "state_code")

p_own <- gw_sequential_map(m_own, "val",
  title = "Own FCIP base rate", legend = "Base rate")
p_lag <- gw_sequential_map(m_lag, "val",
  title = "Neighbour-weighted base rate (gwlag)", legend = "Base rate")
p_own
p_lag
```

The lagged surface is smoother than the raw one: each state now carries the risk
climate of the region around it. Used as a regressor, `base_rate_LM` captures
spillovers — whether farmland is discounted for *regional* risk beyond a state's
own — while `participation_LM` proxies the surrounding insurance uptake.
