---
title: "gwkit example 03 — `estimate_gwss()`: where risk and farmland value co-move"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 5, dpi = 150)
```

Before any regression controls, a descriptive question: *where* is the raw risk
discount visible? `estimate_gwss()` reports **local summary statistics** at each
unit — local means, SDs, and, for two or more variables, the local
covariance/correlation. Mapping the local correlation between the FCIP base rate
(risk proxy) and log farmland value shows where higher assessed risk coincides
with lower land value — the descriptive footprint of the discount that
`estimate_gwr()` later estimates as `a1`.

```{r data}
library(data.table); library(sf); library(ggplot2)
source("_setup.R")     # loads gwkit (dev tree if present) + the ERS framework

data(us_state_ag_census)
d <- data.table::copy(us_state_ag_census)
d[, lv        := log(ag_land_value_per_acre)]
d[, base_rate := fcip_base_rate]
d <- d[is.finite(lv) & is.finite(base_rate) & census_year == max(census_year)]

states <- urbnmapr::get_urbn_map("states", sf = TRUE)
states$state_code <- as.integer(states$state_fips)
states <- states[states$state_code %in% d$state_code, ]
```

```{r gwss}
gwss <- estimate_gwss(
  data = d, variable_list = c("base_rate", "lv"),
  geometry = states, id_col = "state_code",
  distance_metric = "Great Circle", kernel = "gaussian", adaptive = TRUE)

gwss[, state_code := as.integer(state_code)]
# GWmodel names the local correlation column "Corr_<v1>.<v2>"
corr_col <- grep("^Corr", names(gwss), value = TRUE)[1]
corr_col
```

```{r gwss-map}
m <- merge(states, gwss[, .(state_code, local_corr = get(corr_col))],
           by = "state_code")
gw_diverging_map(
  m, "local_corr",
  title    = "Local correlation: FCIP base rate (risk) vs farmland value",
  subtitle = "estimate_gwss(variable_list = c('base_rate','lv'))",
  legend   = "Local\ncorrelation",
  caption  = "Negative (red) = the raw risk discount; bounded [-1, 1]",
  limits   = c(-1, 1))
```

Red regions are where higher assessed risk locally coincides with **lower**
farmland value — the raw discount. This is descriptive and uncontrolled: it
motivates the local regression in example 01 but does not, on its own,
separate risk from the many other things correlated with it.
