---
title: "gwkit example 04 — `gw_consensus_scalar()`: is the risk discount spec-robust?"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 5, dpi = 150)
```

The local risk discount `a1` from example 01 depends on the kernel × distance-
metric choice. `gw_consensus_scalar()` reduces `a1` across **all 50 specifications**
(10 distance presets × 5 kernels) to one robust per-state consensus, plus a
**sign-agreement** share telling us where the discount is stable versus an
artefact of one kernel.

```{r data}
library(data.table); library(sf); library(ggplot2)
source("_setup.R")     # loads gwkit (dev tree if present) + the ERS framework

data(us_state_ag_census)
d <- data.table::copy(us_state_ag_census)
d[, lv            := log(ag_land_value_per_acre)]
d[, base_rate     := fcip_base_rate]
d[, participation := fcip_adoption]
d[, rate_x_part   := base_rate * participation]
d <- d[is.finite(lv) & is.finite(base_rate) & is.finite(participation)]
d_cs <- d[census_year == max(census_year)]

states <- urbnmapr::get_urbn_map("states", sf = TRUE)
states$state_code <- as.integer(states$state_fips)
states <- states[states$state_code %in% d_cs$state_code, ]
```

```{r stack-specs}
specs <- data.table::CJ(dm = gw_distance_metric_names(),
                        kernel = c("gaussian","exponential","bisquare","boxcar","tricube"),
                        sorted = FALSE)

stk <- data.table::rbindlist(lapply(seq_len(nrow(specs)), function(i) {
  r <- tryCatch(estimate_gwr(
        d_cs, unit = "state_code",
        formula = lv ~ base_rate + participation + rate_x_part,
        geometry = states, poly_id = "state_code",
        distance_metric = specs$dm[i], kernel = specs$kernel[i],
        adaptive = TRUE, bw = 12, terms = "base_rate"),
        error = function(e) NULL)
  if (is.null(r)) return(NULL)
  r[term == "base_rate" & estimand == "mean",
    .(state_code = unit_id, distance_metric = specs$dm[i],
      kernel = specs$kernel[i], estimate)]
}), fill = TRUE)

nrow(stk)   # ~ 50 specs x n_states
```

```{r consensus}
cons <- gw_consensus_scalar(
  stk, unit_col = "state_code", value_col = "estimate",
  geometry = states, poly_id = "state_code", agg_fun = stats::median)
cons[, state_code := as.integer(state_code)]
head(cons[, .(state_code, consensus, sign_agreement, n_settings)])
```

```{r consensus-map}
m <- merge(states, cons[, .(state_code, consensus, sign_agreement)],
           by = "state_code")
gw_diverging_map(
  m, "consensus",
  title    = "Spec-robust risk discount (median a1 across 50 specs)",
  subtitle = "gw_consensus_scalar()",
  legend   = "Consensus\na1",
  caption  = "Red = discount robust to kernel/metric choice")
```

```{r agreement-map}
gw_sequential_map(
  m, "sign_agreement",
  title    = "Share of specifications agreeing on the sign of a1",
  subtitle = "gw_consensus_scalar() sign_agreement",
  legend   = "Sign\nagreement")
```

States with high **sign agreement** carry the same-signed discount under nearly
every kernel/metric — the consensus there is trustworthy; low-agreement states
are spec-sensitive and should not be over-interpreted.
