---
title: "gwkit example 06 — building blocks: distance-metric presets & the FE identification problem"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 5, dpi = 150)
```

The estimators are built on a few small exported helpers. This example shows two
of them directly, and uses the second to make the study's central
**identification concern** concrete: separating the risk variable from
cross-sectional fixed effects.

```{r libs}
library(data.table); library(ggplot2)
source("_setup.R")     # loads gwkit (dev tree if present) + the ERS framework
```

## Distance-metric presets

`gw_distance_metric_names()` lists the presets; `resolve_distance_metric()`
returns the `(p, theta, longlat)` triple each one hands to `GWmodel::gw.dist()`.
The spec grids in examples 04–05 iterate over exactly these names.

```{r presets}
gw_distance_metric_names()

presets <- gw_distance_metric_presets()
data.table::rbindlist(lapply(names(presets), function(nm) {
  x <- presets[[nm]]
  data.table(preset = nm, p = x$p, theta = x$theta, longlat = x$longlat)
}))

resolve_distance_metric("Great Circle")          # geodesic
```

## Fixed-effects demeaning and the identification concern

`fixed_effect_model_data_prep()` implements the within transform that
`estimate_gwr(panel=, time=)` uses internally: it drops singleton panels and
replaces each variable by `value − panel_mean + grand_mean`. The point of the
example study is that the **risk discount** `a1` is fragile precisely because the
FCIP base rate varies far more *between* states than *within* them over time — so
demeaning strips most of its variation.

```{r demean}
data(us_state_ag_census)
d <- data.table::copy(us_state_ag_census)
d[, lv        := log(ag_land_value_per_acre)]
d[, base_rate := fcip_base_rate]
d <- d[is.finite(lv) & is.finite(base_rate)]

prep <- fixed_effect_model_data_prep(
  data = d, varlist = c("base_rate", "lv"),
  panel = "state_code", time = "census_year")

prep$NFE                        # number of states with >1 census wave
```

```{r variance-share}
# how much of the base-rate variance is between- vs within-state?
bw_share <- d[, {
  gm <- mean(base_rate, na.rm = TRUE)
  sm <- .SD[, .(mu = mean(base_rate, na.rm = TRUE), n = .N), by = state_code]
  between <- sum(sm$n * (sm$mu - gm)^2, na.rm = TRUE)
  total   <- sum((base_rate - gm)^2, na.rm = TRUE)
  .(between_share = between / total)
}]
bw_share      # closer to 1 => within-state (FE) identification is thin
```

```{r demean-plot}
raw <- d[, .(state_code, census_year, raw = base_rate)]
dm  <- prep$data[, .(state_code, census_year, demeaned = base_rate)]
cmp <- merge(raw, dm, by = c("state_code", "census_year"))

ggplot(cmp, aes(raw, demeaned)) +
  geom_hline(yintercept = mean(cmp$demeaned), colour = "grey70") +
  geom_point(alpha = 0.5, colour = "#00583D") +
  labs(title = "FCIP base rate: raw vs within-state demeaned",
       subtitle = "fixed_effect_model_data_prep() removes each state's level",
       x = "Raw base rate", y = "Within-state demeaned") +
  ers_theme() +
  theme(axis.title.y = element_text(), panel.grid.major.x = element_line(colour = "grey92"))
```

The vertical spread at each raw value is the between-state level the fixed effect
absorbs; the demeaned series — narrow if `between_share` is high — is all a GWFE
`a1` is identified from. That is the caveat flagged in example 01, shown as data:
without meaningful within-state movement in assessed risk, the fixed-effects risk
discount rests on very little.
